Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2021-07-20 paper/paper.tex
Final Tuesday updates/WIP

To enable a simulation it is necessary to make the assumptions, in addition to
those implicit in the description of the model above, that i) All
receptor-ligand interactions are repulsive, as for EphA-ephrinA
coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}; and ii)
all tectal ligand expressions vary exponentially along a given axis, following
Eq.\,\ref{e:tecexp}. The result of running this system are presented in
Fig.\,\ref{f:wt}.
%
Fig.\,\ref{f:wt}A shows a dual-colour map of retinal ganglion cell positions
on the square retina, with red encoding position on the $x$ axis (temporal to
nasal) and green encoding position on the $y$ axis (ventral to
dorsal). Fig.\,\ref{f:wt}B presents a fishnet plot of the expected location on
the tectal surface of each RGC at the end of the development process. Thus,
the `red' RGC from the ventro-nasal corner should map to the caudo-medial
corner of the tectum and the `green' RGC from the dorso-temporal corner should
find its way to the rostro-lateral corner on the tectum. The lines between the
dots of Fig.\,\ref{f:wt}B indicate the adjacency relationships between
cells. Thus, colour in each panel indicates a branch's location of origin on
the retina.
%
Fig.\,\ref{f:wt}C shows the initial state of the system, with branches
stochastically arranged around the rostral edge of the tectum. Each dot
represents the centroid of $n=8$ branches per RGC axon. Figs.\,\ref{f:wt}D-E
show the development of the system with time, with Fig.\,\ref{f:wt}E showing

\subsection*{Chemoaffinity---Wildtype}

>>>>

To enable a simulation it is necessary to make the assumptions (in addition to
those implicit in the description of the model above) that i) All
receptor-ligand interactions are repulsive, as for EphA-ephrinA
coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}; and ii)
all tectal ligand expressions vary exponentially along a given axis, following
Eq.\,\ref{e:tecexp}. The result of running this system are presented in
Fig.\,\ref{f:wt}.
%
Fig.\,\ref{f:wt}Ai shows a dual-colour map of retinal ganglion cell positions
on the square retina, with red encoding position on the $x$ axis (temporal to
nasal) and green encoding position on the $y$ axis (ventral to
dorsal). Fig.\,\ref{f:wt}Aii presents a fishnet plot of the expected location on
the tectal surface of each RGC at the end of the development process. Thus,
the `red' RGC from the ventro-nasal corner should map to the caudo-medial
corner of the tectum and the `green' RGC from the dorso-temporal corner should
find its way to the rostro-lateral corner on the tectum. The lines between the
dots of Fig.\,\ref{f:wt}Aii indicate the adjacency relationships between
cells. Thus, colour in each panel indicates a branch or axon's location of
origin on the retina.
%
Fig.\,\ref{f:wt}B shows the initial state of the system, with branches
stochastically arranged around the rostral edge of the tectum. Each dot
represents the centroid of $n=8$ branches per RGC axon. Figs.\,\ref{f:wt}C and \ref{f:wt}D
show the development of the system with time, with Fig.\,\ref{f:wt}D showing a
well ordered arrangement, with a majority of adjacency relationships matching
the experimental prediction (i.e. there are few crossed lines). The
intermediate timepoint at $t=40$ indicates that branches quickly cover the simulated
tectal surface, with final adjacency relationships developing over a longer
timescale. This dynamic behaviour is also evident in Fig.\,\ref{f:wt}E, which
shows \textbf{RMS error or some metric} of the axon centroids with respect to
the experimental prediction. Fig.\,\ref{f:wt}F shows the final
location of each branch in the simulation, which gives a more continuous view
of the order present in Fig.\,\ref{f:wt}D. Finally, Fig.\,\ref{f:wt}G shows the
branches and their centroid path history for five selected RGCs, with path
histories strongly resembling those presented by \citet{simpson_simple_2011}.

This result shows that a gradient following model based on receptor-ligand
interactions and incorporating axon-axon repulsion is able to generate the
basic mapping observed in the retinotectal projection.

\subsection*{Manipulations}

To determine if the model can explain the many surgical and genetic
manipulations described in the literature
\citep*[for a review, see][]{goodhill_retinotectal_1999} we performed
simulations matching those in \citet{simpson_simple_2011} where we either
rotated or translocated grafts of the tectum, ablated portions of the retina
and/or the tectum or manipulated the receptor or ligand expressions in the
model...

\subsection*{Grating simulations}

I'd like to do this if it's not too much work!

\subsection*{Is chemoaffinity alone sufficient to explain the retinotectal projection?}
